---
title: "Quickstart"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quickstart}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(vdbR)
library(dplyr)
library(ggplot2)
library(phyloseq)
library(microViz)
theme_set(theme_classic())


DEBUG = FALSE
connect_database(bundled=Sys.getenv("GITHUB_ACTIONS") != "")
get_table_from_database("frozen_set_ag")


```

Lets start by getting some metadata. For simplicity we will pull a frozen set, but in practice it could be any dataframe containing sampleids. We will use phyloseq's `tax_glom` to aggregate the data at the genus level.

```{r}
set.seed(123)
phy_amplicon <- vdb_make_phylo(frozen_set_ag %>% dplyr::sample_n(50)) %>% 
  phyloseq::tax_glom("genus")  %>% 
  microViz::tax_fix()
```
Lets add in the alpha diversity:
```{r}
sample_data(phy_amplicon) <- cbind(
  sample_data(phy_amplicon),
  suppressWarnings(estimate_richness(phy_amplicon, measures = "InvSimpson"))
)
ggplot(sample_data(phy_amplicon) %>% as_tibble() %>% filter(day_relative_to_hct > -10 & day_relative_to_hct < 30), aes(x=day_relative_to_hct, y=InvSimpson, color=institution)) + geom_smooth() + geom_point()
```



The MicroVis takes a lot of the annoying considerations out of building the classic stacked bar plots.
```{r}
tax_table(phy_amplicon) <- clean_refseq_genus(phy_amplicon)

mypal <- make_microviz_palette(phy_amplicon, n=30, rank="genus", shuf_genus = F) 
phy_amplicon %>%
  comp_barplot(tax_level = "genus", palette = mypal, n_taxa=30, tax_order = names(mypal)) +
  facet_grid( ~institution, scales = "free", space = "free") + 
  theme(axis.text.x = element_blank()) + 
  theme(legend.position = "bottom", legend.direction = "horizontal")

```

The palette used in previous publications utilized higher-order colors as well.  Recreate this behavior by setting `n` to the number of genera as follows:

```{r}
fullpal <- make_microviz_palette(phy_amplicon, n=ntaxa(phy_amplicon), rank="genus", shuf_genus = FALSE) 


(p_comp <- phy_amplicon %>%
  comp_barplot(tax_level = "genus", palette = fullpal, n_taxa=ntaxa(phy_amplicon), bar_outline_colour = NA, tax_order = names(fullpal) ) +
  theme(axis.text.x = element_blank()) + 
  theme(legend.position = "bottom", legend.direction = "horizontal") + 
  guides(fill="none") + 
  scale_y_continuous(expand=c(0, 0)) )


```



Lets plot some beta diversity
```{r}
dists <- distance(phy_amplicon, method = "bray")
ord <- ordinate(phy_amplicon, method = "PCoA", distance = dists)
plot_ordination(phy_amplicon, ord, color="institution")

```

```{r}
tsne <- make_tsne(dists, perplexity = 10) %>% 
  left_join(data.frame(sample_data(phy_amplicon)) %>% tibble::rownames_to_column("sampleid"))

ggplot(tsne, aes(t1_scaled, t2_scaled, color = institution)) +
  geom_point() 
```

